Authors' response to 3rd round of comments on BMED-D-20-02413

### 2021-01-09 authors_response3_BMED-D-20-02413.Rmd
updated shielding paper, fixed trainmodel.R

These tables were reviewed for statistical disclosure control not only by the authors but also by officials of Public Health Scotland, our national public health agency.  We guess that the reviewer's concerns relate to frequency tables containing small numbers. For instance Table 1 shows that one of 483 individuals in Scotland who entered critical care for COVID-19 and survived had been diagnosed with liver disease.  If an attacker already knew the identity of an individual who had been diagnosed with liver disease and had entered critical care for COVID-19 and survived, they would be able to infer a match of this individual to this cell, but this would not provide the attacker with any new information about this individual.

@@

For the REACT-SCOT study, we specified in advance a stringent definition of severe COVID-19 so that ascertainment of severe cases would not be affected by triage.  This was based on advice from a senior critical care specialist (Dr Lone, co-author of this study).  As our case definition  included all deaths certified with COVID-19 as underlying cause, triage of individuals who were assessed (on the basis of age and underlying conditions) as unlikely to benefit from critical care would have been ascertained through their death certificates.

@@

The design of this study is based on incidence density sampling in which matched controls were sampled from those who had not yet tested positive for COVID-19 on the day that the case was diagnosed. This ensures that the conditional odds ratios are interpretable as rate ratios. For each date that one or more cases were diagnosed, all those in the national population register who had tested positive up to that date were excluded from the sampling of controls. This is correct, on the assumption that those who have already tested positive are not at risk of severe disease.

@@

The rate ratio of 4.1 associated with antipsychotic agents is considerably larger than the rate ratios associated with heart disease and most of the other listed conditions.

@@

As the caption states, this table was filtered to show only drug classes for which associations are significant at p<0.001.  We have added the rate ratio for SSRIs to the text

@@

This is provided with the PLOS Medicine paper.

>>>>

These tables were reviewed for statistical disclosure control by officials of Public Health Scotland, our national public health agency.  We guess that the reviewer's concerns relate to frequency tables containing small numbers. For instance Table 1 shows that one of 483 individuals in Scotland who entered critical care for COVID-19 and survived had been diagnosed with liver disease.  **This does not enable identification of an individual**. If an attacker already knew the identity of an individual who had been diagnosed with liver disease and had entered critical care for COVID-19 and survived, they would be able to infer a match of this individual to this cell, but this would not provide the attacker with any new information about this individual.

@@

The definition of severe COVID-19 used for this study ensures that case ascertainment is not affected by triage of those assessed as unlikely (on the basis of age and underlying conditions) to benefit from critical care.  As our case definition included all deaths certified with COVID-19 as underlying cause, these cases would have been ascertained through their death certificates.

We have added a sentence (highlighted) to the first paragraph of the Methods section to make this point clear.

@@

This study was based on incidence density sampling in which matched controls were sampled from those who had not yet tested positive for COVID-19 on the day that the case was diagnosed. With this design, the conditional odds ratios are interpretable as rate ratios (ratios of hazard rates) in the population at risk.  (Greenland S, Thomas DC. On the need for the rare disease assumption in case-control studies. American Journal of Epidemiology 1982;116:547–553,  https://doi.org/10.1093/oxfordjournals.aje.a113439)

For each date that one or more cases were diagnosed, all those in the national population register who had tested positive up to that date were excluded from the sampling of controls.  This defines the population at risk of severe COVID-19 as all those who have not yet tested positive.  For clinicians and public health officials this is the relevant definition of the population at risk, as population-wide screening of immunity status has not been conducted.

We have added a sentence (highlighted) to the second paragraph of the Methods section to explain this technical point.

@@

We have added the following sentence (highlighted) to the Discussion

As the rate ratio of 4.1 associated with use of antipsychotic agents is larger than the rate ratios of about 2 associated with common risk conditons such as Type 2 diabetes and heart disease that we have previously reported from  this case-control study, it is unlikely that underdiagnosis of these risk conditions in users of antipsychotic agents can explain this association.

@@

The rate ratio associated with SSRIs is given in Table 5, which includes all the drug classes listed by Laporte and Healy.

SSRIs do not appear in Table 4 because (as the caption states) this table was restricted to those without listed conditions and filtered to show only drug classes for which associations are significant at p<0.001.

@@

The ICD-10 and BNF drug codes are available as a linked attachment on the ENCEPP registration page.  The link is at Section 19 "Other relevant documents":

http://www.encepp.eu/encepp/openAttachment/documents.otherDocument-0/35565

This link has been added to the manuscript under "Availability of data and materials", and the ENCEPP registration has been updated to cite the PLOS Medicine paper as the source for the full study protocol.
